The five pillars of computational reproducibility

### 2023-05-03 docs/5pillars.Rmd
working on 5 pillars

Computational reproducibility is in theory a simple premise, however in practice it is difficult to
achieve.
Here, we present a framework to improve reproducibility, called the five pillars.
These are a combination of best practices which ensure that the work can be reproduced quickly and
easily, and will endure relatively long time-frames.

@@

Computational reproducibility is the ability to use the data and code provided by a research study to
regenerate the outputs including figures and tables to confirm the study's findings.
Reproducibility is only the first step towards overall reliability.
Replicability is the ability to repeat the entire study and come up with similar overall results.
Robustness refers to the ability for the findings to be generalisable to other situations
@Nosek2022-zc.

If a study is irreproducible or unreplicable, then it is of little value.

>>>>

Computational reproducibility is in theory a simple premise, however in practice it is difficult to
achieve.
Here, we present a framework to improve reproducibility, called the five pillars.
These are a combination of best practices which ensure that the computational research work can be
reproduced quickly and easily, and will endure relatively long time-frames.

@@

In research, Computational reproducibility is the ability to use the data and code provided by a
study to regenerate the outputs including figures and tables to confirm the study's findings.
Reproducibility is only the first step towards overall reliability, for example a study may be
reproducible, but suffer from analytical problems that invalidate the results.
If a study is reproducible, then at least these analytical issues can be more easily identified and
rectified.
So reproducibility is a necessary stepping-stone that in theory should enhance the overall reliability
of computational research including replicability and robustness.
Replicability being defined as the ability to repeat the entire study and come up with similar overall
results.
Robustness refers to the ability for the findings to be generalisable to other situations
@Nosek2022-zc.

Unreliable research is a waste of resources and time, and can mislead readers to pursue doomed
research directions.
Much has been written on how irreproducibility has reached crisis levels in psychology and life
sciences [@Baker2015-bs; @Begley2012-di].
Indeed in bioinformatics, the situation is grim with only a very fraction of studies
being able to be reproduced @Ioannidis2009-hv, which brings into question the reliability of those
studies.
Irreproducible research has the potential to place patient safety at risk, as demonstrated by an
earlier case study @Baggerly2010-nr.

As computational researchers, we are theoretically able to achieve 100% analytical reproducibility,
however in practice this is rarely achieved.
Data and code availability, ability to install software easily, and other factors mean that most
computational research works are unable to be reproduced.

A number of guides recommending enhanced computational reproducibility practices have emerged
[@Peng2011-ia; @Tan2010-is; @Sandve2013-lk; @Piccolo2016-ji; @Gruning2018-sg; @Lewis2016-qh;
@Brito2020-ld; @Akalin2018-eg ]
and this article attempts to unify these concepts and assemble a group of best practices where
computational reproducibility can be considered as complete.

```{r,5pillars}

plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(0, 10), ylim=c(0, 10))
polygon(c(0,1.8,1.8,0),c(1,1,6,6))
polygon(c(2,3.8,3.8,2),c(1,1,6,6))
polygon(c(4,5.8,5.8,4),c(1,1,6,6))
polygon(c(6,7.8,7.8,6),c(1,1,6,6))
polygon(c(8,9.8,9.8,8),c(1,1,6,6))
polygon(c(-0.2,10,10,-0.2),c(-0.4,-0.4,0.8,0.8))
polygon(c(-0.2,5,10),c(6.5,8.5,6.5))
